---
title: "Chapter 2 - Descriptive Analytics"
author: "Gementiza"
date: "Canvas Group 2"
output:
  html_document: default
---

## Plotly and Dash

**Plotly**: visualization library

**Dash**: framework for interactive dashboards


## Histogram vs. Density Curve

```{r}

library(ggplot2)
library(plotly)

set.seed(123)
x <- rnorm(1000, mean = 50, sd = 10)
p <- ggplot(data.frame(x), aes(x)) +
geom_histogram(aes(y = ..density..),
bins = 30,
fill = "green",
color = "black") +
geom_density(color = "gold", linewidth = 1) +
labs(title = "Probability Distribution")

ggplotly(p)

```

## Box and Whisker Plots
Find the median, IQR, whiskers, and outliers.

```{r}

set.seed(123)

data <- data.frame(
Group = rep(c("A", "B"), each = 50),
Score = c(rnorm(50, 75, 5), rnorm(50, 85, 7))
)
p <- ggplot(data, aes(Group, Score, fill = Group)) +
geom_boxplot() +
labs(title = "Box and Whisker Plot")

ggplotly(p)
```


## Scatter Plots
Discuss relationships, correlations, and trends

```{r}

df <- data.frame(
StudyHours = rnorm(100, 1, 10),
Score = 50 + 5 * rnorm(100, 1, 10))
p <- ggplot(df, aes(StudyHours, Score)) +
geom_point(alpha = 0.7) +
geom_smooth(method = "lm", se = TRUE) +
labs(title = "Scatter Plot with Trend Line")

ggplotly(p)

```

## Time Series Plots
Highlight trend, seasonality, and irregular components

```{r}

time_data <- data.frame(
Date = seq.Date(from = as.Date("2024-01-01"),
by = "month",
length.out = 12),
Sales = cumsum(runif(12, 10, 30))
)
p <- ggplot(time_data, aes(Date, Sales)) +
geom_line() +
geom_point() +
labs(title = "Time Series Plot")

ggplotly(p)

```

## Bar Charts
For categorical data

```{r}

category_data <- data.frame(
Category = c("Math", "Science", "English"),
Count = c(40, 55, 30)
)
p <- ggplot(category_data, aes(Category, Count)) +
geom_bar(stat = "identity", fill = "steelblue") +
labs(title = "Bar Chart")

ggplotly(p)

```

## Correlation Heatmaps
Explain correlation strength and direction

```{r}

library(corrplot)

num_data <- mtcars[, c("mpg", "hp", "wt", "disp")]
corr_matrix <- cor(num_data)
corrplot(corr_matrix, method = "color",
addCoef.col = "white")

```

## Geographical Mapping
Demographic, sales, and logistics

```{r}

library(leaflet)

leaflet() %>%
addTiles() %>%
addMarkers(
lng = 123.8854,
lat = 10.3157,
popup = "Cebu City"
)


```